---
title: "R Notebook"
output: html_notebook
---

#I used two files here: one was from "Table 10. Supply and disposition of electricity, 1990 through 2023" and the other was "Table 5. Electric power industry generation by primary energy source, 1990 through 2023," the latter of which was mainly used to visualize how generation overall has actually declined, which helped contexualize the data on imports. Both were from the same EIA SEP Tables for IN dataset and were sourced from Form EIA-923, Power Plant Operations Report and predecessor forms, Annual Electric Generator Report, Form EIA-861, Annual Electric Power Industry Report and Form EIA-111, Quarterly Imports and Exports Report. Link: https://indiana-my.sharepoint.com/:x:/g/personal/marnmead_iu_edu/EWVVMOXz4HJJm-jPHmacmUAByQvQE3EoAU6QTWqrMIpIJw?e=ygOc1i


```{r}
df_sales <- read.csv('Book 11(Sheet1).csv')
```


```{r}
df_imports <- read.csv('imports(Sheet1).csv')
```


```{r}
library(dplyr)
library(tidyverse)
library(stringr)
```
#I need to flip the data - make the years a column and make the category column a set of rows

```{r}
df_long <- df_imports %>%
  pivot_longer(cols = starts_with("Year."),
               names_to = "Year",
               values_to = "Value") %>%
  mutate(Year = sub("Year\\.", "", Year)) %>% 
  pivot_wider(names_from = Category, values_from = Value)

```

#Now I need to keep only the columns I'm analyzing, which are the various categories where energy came from, how much of it was sold, and whether it was used or lost.


```{r}
df_imports_clean <- df_long %>%
select(Year, `Total net generation`, `Net interstate imports`, `Total supply`, `Unaccounted`, `Estimated losses`, `Total electric industry sales`, `Direct use`)
```

#I noticed the commas were preventing me from changing to numeric so I removed them

```{r}
df_imports_clean$Year <- as.factor(df_imports_clean$Year)
```
```{r}
df_no_commas <- df_imports_clean %>%
  mutate(`Total net generation` = str_remove_all(`Total net generation`, ",")) %>%
mutate(`Net interstate imports` = str_remove_all(`Net interstate imports`, ",")) %>%
 mutate (`Total supply` = str_remove_all(`Total supply`, ",")) %>%
mutate  (`Unaccounted` = str_remove_all(`Unaccounted`, ",")) %>%
mutate  (`Estimated losses` = str_remove_all(`Estimated losses`, ",")) %>%
  mutate(`Direct use` = str_remove_all(`Direct use`, ","))

```


```{r}
df_imports_clean <- df_no_commas %>%
  mutate(`Total electric industry sales` = str_remove_all(`Total electric industry sales`, ","))
```

#Now making some key columns numeric

```{r}
df_imports_clean$`Total electric industry sales` <- as.numeric(as.character(df_imports_clean$`Total electric industry sales`))
```

```{r}
df_imports_clean$`Net interstate imports` <- as.numeric(as.character(df_imports_clean$`Net interstate imports`))
```
```{r}
df_imports_clean$`Total supply` <- as.numeric(as.character(df_imports_clean$`Total supply`))
```

#Then I wanted to create a column that gave me the ratio of interstate imports to total supply. Since utilities tend to purchase power when they are unsure they have enough to meet demand, a higher ratio may indicate greater struggle to meet demand

```{r}
df_imports_2 <- df_imports_clean %>%
  mutate(ratio = `Net interstate imports`/`Total supply`)

```

#Then I plotted it. There appears to be a dramatic increase

```{r}
ggplot(df_imports_2, aes(x = Year, y = ratio)) + geom_line(size =1) + geom_point(size = 2) + labs(title = "Ratio of Imports to Total Generation Over Time", x = "Year", y = "Ratio") + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
#I also decided to plot electricity sales to see if increased demand was driving the struggle to meet demand. It appears sales have actually declined, meaning the real culprit may lie in power generation methods themselves (likely the retiring of coal plants)

```{r}
ggplot(df_imports_clean, aes(x = Year, y = `Total electric industry sales`)) + geom_line(size =1) + geom_point(size = 2) + labs(title = "Indiana Electricity Sales Over Time", x = "Year", y = "Sales") + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}
write.csv(df_imports_clean, 'electricity_data_clean.csv')
```

```{r}
write.csv(df_imports_2, 'import_data.csv')
```

#Flourish graphic: https://public.flourish.studio/visualisation/22795170/

