---
title: "R Notebook"
output: html_notebook
---
#The data I'm working with is from a set of files from Energy Information Administration called "SEP Tables for IN": https://indiana-my.sharepoint.com/:x:/g/personal/marnmead_iu_edu/EWVVMOXz4HJJm-jPHmacmUAByQvQE3EoAU6QTWqrMIpIJw?e=HSq3Yf. The specific file I'm using is "Table 5. Electric power industry generation by primary energy source, 1990 through 2023" which sources from U.S. Energy Information Administration, Form EIA-923, Power Plant Operations Report and predecessor forms.																												
```{r}
df <- read.csv('generation(5.csv')
```

```{r}
library(dplyr)
library(stringr)
library(tidyverse)
```

#I'm going to remove the sub columns so I only get the total for each category. I'm also removing the "other gas" category because it doesn't appear to be a significant source of energy. It's missing from the last decade of data, so I assume it's not used anymore.

```{r}
df2 <- df[c(2,3,4,5,10,11,13,18),]
```

#Next I'm selecting out the percent share columns. I had to do this one by one for some reason, perhaps because I chose to overwrite the dataframe rather than creating a new one. 

```{r}
df2 <- df2 %>%
select(-Percent.share.2000, Percent.share.2010, Percent.share.2020, Percent.share.2023)
```

#I want the years to be in one column

```{r}
df_long <- df2 %>%
  pivot_longer(cols = starts_with("Year."),
               names_to = "Year",
               values_to = "Value") %>%
  mutate(Year = sub("Year\\.", "", Year))

```

#Then I'll pivot_wider to make each category row a column.

```{r}
df_clean <- df_long %>%
    pivot_wider(names_from = X, values_from = Value)
```


```{r}
write.csv(df_clean, 'generation_by_source.csv')
```

#Flourish: https://public.flourish.studio/visualisation/22526856/
